NumPy---Array Computation for Python

### 2020-01-04 body.tex
Rework intro / summary

%NumPy is the fundamental array computation library for the Python
%ecosystem \cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
With its first stable release in October 2006, NumPy unified
the scientific Python community and now underpins almost every library
that does numerical computation, including SciPy\cite{virtanen2019scipy},
pandas\cite{mckinney-proc-scipy-2010}, scikit-learn\cite{pedregosa2011scikit},
and scikit-image\cite{vanderwalt2014scikit}.
Not surprisingly, many important scientific research projects and
results crucially rely on NumPy.
It plays an essential role in many research analysis pipelines in
fields as diverse as physics, biology, astronomy, neuroscience,
material science, engineering, and chemistry.
% w/ citations
In particular, NumPy played a fundamental role in discovering
gravitational waves\cite{pycbc} and the first imaging of a black
hole\cite{eht-imaging}, and lets us explore our
universe\cite{jenness2018lsst}.

@@

NumPy consists of the \code{ndarray}-object along with utility functions that
operate on it.
Because of its inherent simplicity—being a pointer to memory with some
associated meta-data about shape, data-type, and so forth—the NumPy array is
the {\it de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
{\it Application Programming Interface} (API) has become ubiquitous as
a language for tensor computation---witnessed by its use in popular
deep learning libraries such as PyTorch\cite{pytorch}.

@@

The \code{ndarray} data structure stores regularly spaced elements of a single
data type in a contiguous block of memory; along with meta-data such as shape,
it allows for the efficient representation of $n$-dimensional arrays.
\fixme{discuss Figure~\ref{fig:basic-functionality} .. vectors / matrices / and then more ...}
More details about the data structure are given in ``The NumPy array:
a structure for efficient numerical computation.''\cite{vanderwalt2011numpy}.

>>>>

%NumPy is the fundamental array computation library for the Python
%ecosystem \cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
%With its first stable release in October 2006, NumPy unified
%the scientific Python community and now underpins almost every library
%that does numerical computation, including SciPy\cite{virtanen2019scipy},
%pandas\cite{mckinney-proc-scipy-2010}, scikit-learn\cite{pedregosa2011scikit},
%and scikit-image\cite{vanderwalt2014scikit}.
%Not surprisingly, many important scientific research projects and
%results crucially rely on NumPy.
%It plays an essential role in many research analysis pipelines in
%fields as diverse as physics, biology, astronomy, neuroscience,
%material science, engineering, and chemistry.
%% w/ citations
%In particular, NumPy played a fundamental role in discovering
%gravitational waves\cite{pycbc} and the first imaging of a black
%hole\cite{eht-imaging}, and lets us explore our
%universe\cite{jenness2018lsst}.

@@

With its first stable release in October 2006, NumPy unified
the scientific Python community and now underpins almost every library
that does numerical computation, including SciPy\cite{virtanen2019scipy},
pandas\cite{mckinney-proc-scipy-2010}, scikit-learn\cite{pedregosa2011scikit},
and scikit-image\cite{vanderwalt2014scikit}.
NumPy consists of the \code{ndarray}-object along with utility functions that
operate on it.
Because of its inherent simplicity—being a pointer to memory with some
associated meta-data about shape, data-type, and so forth—the NumPy array is
the {\it de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
{\it Application Programming Interface} (API) has become ubiquitous as
a language for tensor computation---witnessed by its use in popular
deep learning libraries such as PyTorch\cite{pytorch}.

@@

The \code{ndarray} data structure stores regularly spaced elements of a single
data type in a contiguous block of memory; along with meta-data such as shape,
it allows for the efficient representation of $n$-dimensional arrays.

\fixme{discuss Figure~\ref{fig:basic-functionality} .. vectors / matrices / and then more ...
indexing and slicing}

More details about the data structure are given in ``The NumPy array:
a structure for efficient numerical computation.''\cite{vanderwalt2011numpy}.

### 2020-01-04 summary.tex
Rework intro / summary

NumPy is the fundamental array computation library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
It has been actively developed for over two decades and has
more than 850 contributors, nearly 300,000 dependent repositories
on GitHub, and millions of downloads per year.
It is extensively used in research and teaching.
According to a 2018 survey, NumPy is the second most commonly used machine
learning tool among enterprises \cite{451report2018}.
NumPy continues to evolve to meet the needs of its community, to
remain relevant in an ever-changing computing landscape, and to be a
fundamental pillar of support for the next decade of data science in
Python.
In this article, we provide an overview of the project and library
organization,  highlight some recent technical developments, and
outline our future plans.

>>>>

% - 2-3 sentences of basic-level introduction to the field;
Software libraries for array computation provide powerful notation
for accessing, manipulating, and operating on aggregate data \cite{iverson1980notation}.
% - a brief account of the background and rationale of the work;
NumPy is the fundamental array computation library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
Not surprisingly, many important scientific research projects and results
crucially rely on NumPy.
It plays an essential role in many research analysis pipelines in fields as
diverse as physics, biology, astronomy, neuroscience, material science,
engineering, and chemistry.
In particular, NumPy played a fundamental role in discovering gravitational
waves\cite{pycbc} and the first imaging of a black hole\cite{eht-imaging}, and
lets us explore our universe\cite{jenness2018lsst}.
It is one of the most commonly used machine learning tools among enterprises
\cite{451report2018}.
% - a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
Here we provide an overview of the project and library organization,  highlight
some recent technical developments, and outline our future plans.
% - 2-3 sentences putting the main findings into general context so it is
%   clear how the results described in the paper have moved the field forwards.
NumPy continues to evolve to meet the needs of its community, to remain
relevant in an ever-changing computing landscape, and to be a fundamental
pillar of support for the next decade of data science in Python.
